White Dwarf Mergers on Adaptive Meshes\\ II. Collisions

### 2017-09-09 papers/wdmerger_II/paper.tex
Update discussion of iso7

\isoseven\ also has the peculiar property that after the detonation had passed through the WDs,
the region at the central contact point did not fully stop burning, yielding a quasistatic energy
release that added a couple of million extra timesteps to the simulation (in addition to the few
hundred it normally takes) before we manually terminated the run. The full effect of these additional
timesteps only amounted to a $10^{-4}$ relative change in the nickel production. By examining the behavior
of zones near the collision point, we discovered the reason for this. The zones in question have a composition that
is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
step following the advective update. This then causes the burning timestep limiter to kick in and make the
timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.

>>>>

\isoseven\ also has the peculiar property that after the detonation has passed through the WDs,
the region at the central contact point does not fully stop burning, yielding a quasistatic energy
release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
hundred it normally takes). The full effect of these additional timesteps only amounts to a trace change
in the nickel production. By examining the behavior of zones near the collision point, we discovered the
reason for this. The zones in question have a composition that
is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
step following the advective update. This then causes the burning timestep limiter to kick in and make the
timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.
